---
title: "Parallel Institutions and Civil Resistance"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

What is the impact of parallel institution-building on non-state actors’ strategic campaigns? Non-state campaigns, both violent and nonviolent, occasionally attempt to develop independent institutions for the provision of public goods and services and for the execution of political tasks. In this paper, I employ the Nonviolent and Violent Campaign Outcomes 2.1 (NAVCO) dataset to measure how building parallel institutions affects the longevity and success rates of campaigns. Using survival analysis, I show that education and social welfare institutions can help campaigns achieve their goals, but that this effect diverges based on the violence of a campaign.

* [Hypotheses](#Hypotheses)
* [DVs and IVs](#Dependent-and-Independent-Variables)

### Hypotheses

* **Hypothesis 1**: Social welfare and educational institutions help political resistance campaigns achieve their goals.
* **Hypothesis 2**: Campaigns with traditional or new media systems last longer and have higher rates of success than those without.
* **Hypothesis 3**: Violent campaigns are benefitted by building education systems, while nonviolent campaigns are benefitted by social welfare systems.

### Dependent and Independent Variables
**DV**: Campaign length; campaign success or failure.

**IV**: Parallel institutions:

* Law Enforcement
* Education
* Social Welfare
* Traditional Media
* New Media
* Courts

Condition: 

* Violence or nonviolence of campaign

### Methods

* Random Effects Linear Regression: Campaign Length vs. Parallel Institutions
* Binomial Logistic Regression: Campaign Success vs. Parallel Institutions
* Multinomial Logistic Regression: Campaign Success OR Failure vs. Parallel Institutions
* Competing Risks Survival Analysis: Campaign Success Hazard Rates on Parallel Institutions, with Failure as Competing Event

### Variable Characteristics

```{r}
library(tidyverse)
library(haven)
colonies <- read_csv("COLDAT_colonies.csv")
df <- haven::read_dta("NAVCO2-1_ForPublication.dta") 
gdp <- readxl::read_xlsx("mpd2020.xlsx", sheet = "Full data" )

gdp_wb <- read_csv("worldbank_gdp.csv") %>% 
  select(Year, PNG_2011, MDV_2011) %>% 
  pivot_longer(cols = c(PNG_2011, MDV_2011), names_to = "country", values_to = "gdppc") %>% 
  mutate(loc_cow = case_when(country == "PNG_2011" ~ 910,
                             country == "MDV_2011" ~ 781)) %>% 
  select(-country, year = Year)


df <- df %>% 
  mutate(across(starts_with("pi_"), ~ case_when(. >= 0 ~ .,
                                                . < 0 ~ NA)))
df <- df %>% 
  mutate(in_media = na_if(in_media, -99)) %>% 
  mutate(repression = na_if(repression, -99)) %>% 
  mutate(camp_support = na_if(camp_support, -99)) %>% 
  mutate(camp_size = na_if(camp_size, -99)) %>% 
  mutate(regime_support = na_if(regime_support, -99)) %>% 
  mutate(indiscrim = na_if(indiscrim, -99)) %>% 
  mutate(prim_meth = na_if(prim_meth, -99)) %>% 
  mutate(sec_defect = na_if(sec_defect, -99)) %>% 
  mutate(sdirect = na_if(sdirect, -99)) %>% 
  mutate(prim_meth = case_when(prim_meth == 0 ~ 1,
                               TRUE ~ 0))

df <- df %>% mutate(cyearplus = cyear + 1)

df %>% 
  group_by(id) %>% 
  mutate(final_year = case_when(cyear == max(cyear) ~ 1,
                                TRUE ~ 0)) %>% 
  ungroup() -> df

df <- df %>% 
  group_by(id) %>% 
  mutate(years_active = row_number()) 

df <- df %>%
  mutate(years_plus = years_active + 1)

df <- df %>% 
  mutate(end_status = case_when(success == 1 ~ 2,
                                (success == 0 & final_year == 1) ~ 1,
                                TRUE ~ 0)) 


gdp <- gdp %>%
  mutate(loc_cow = countrycode::countrycode(countrycode, origin = "iso3c", destination = "cown"))

gdp <- gdp %>% 
  mutate(loc_cow = case_when(countrycode == "CSK" ~ 315,
                             countrycode == "HKG" ~ 997,
                             countrycode == "PSE" ~ 666,
                             countrycode == "SRB" ~ 345,
                             countrycode == "SUN" ~ 365,
                             countrycode == "YUG" ~ 345,
                             TRUE ~ loc_cow))

gdp <- gdp %>% 
  select(loc_cow, year, gdppc)

gdp <- gdp %>% 
  group_by(loc_cow, year) %>% 
  summarize(gdppc = sum(gdppc))

df <- df %>% 
  left_join(gdp, by = c("loc_cow", "year"))



colonies %>% 
  pivot_longer(cols = ends_with("_max"), names_to = "colonizer", values_to = "start_year", names_repair = "unique") %>% 
  select(country, colonizer, start_year) %>% 
  arrange(country, colonizer) %>% 
  filter(!is.na(start_year)) -> colonies_long


colonies_long %>% 
  rename(year = start_year) %>%
  mutate(colonizer_name = str_extract(colonizer, "(?<=[a-z]{1,16}\\.).*(?=\\_max)")) %>%
  mutate(start_year = case_when(grepl("colstart", colonizer) ~ year)) %>% 
  mutate(end_year = case_when(grepl("colend", colonizer) ~ year)) %>% 
  group_by(country, colonizer_name) %>% 
  summarize(start_year = mean(start_year, na.rm = TRUE), end_year = mean(end_year, na.rm = TRUE)) -> colonies_long


colonies_long$year <- mapply(seq,colonies_long$start_year,colonies_long$end_year,SIMPLIFY=FALSE)

colonies_long %>% 
  unnest(year) %>% 
  select(-start_year, -end_year) %>% 
  mutate(colony = countrycode::countrycode(colonizer_name, origin = "country.name", destination = "cown")) -> colonies_long 

colonies_long <- colonies_long %>% 
  mutate(cown = countrycode::countrycode(country, origin = "country.name", destination = "cown")) 

df <- df %>% 
  left_join(colonies_long %>% select(-country, -colonizer_name), by = c("year" = "year", "loc_cow" = "cown"))

df <- df %>% 
  mutate(colony = case_when(is.na(colony) ~ 0,
                            TRUE ~ colony)) 





colony_boolean <- colonies_long %>% 
  distinct(cown) %>% 
  mutate(colony_bool = TRUE)

df <- df %>% 
  left_join(colony_boolean %>% select(cown, colony_bool), by = c("loc_cow" = "cown")) %>% 
  mutate(colony_bool = case_when(is.na(colony_bool) ~ FALSE,
                                 TRUE ~ colony_bool))

polity <- readxl::read_xls("p5v2018.xls")

polity <- polity %>% 
  mutate(polity = case_when(polity < -10 ~ NA,
                            TRUE ~ polity))

polity <- polity %>% 
  mutate(cown = countrycode::countrycode(country, origin = "country.name", destination = "cown")) %>%
  mutate(cown = case_when(grepl("serbia", country, ignore.case = TRUE) ~ 345,
                          TRUE ~ cown))



xrcomp <- polity %>% 
  select(country, year, xrcomp) %>% 
  mutate(cown = countrycode::countrycode(country, origin = "country.name", destination = "cown")) %>%
  mutate(cown = case_when(grepl("serbia", country, ignore.case = TRUE) ~ 345,
                          TRUE ~ cown)) %>% 
  select(cown, year, xrcomp)

df <- df %>% 
  left_join(xrcomp, by = c("loc_cow" = "cown", "year"="year"))

df <- df %>% distinct() 

df <- df %>% 
  mutate(colony = factor(colony, ordered = FALSE)) %>% 
  mutate(colony = relevel(colony, ref = "0"))
```

```{r}

agg_df <- df %>% 
  group_by(camp_name) %>% 
  summarise(end_status = max(end_status), loc_cow = max(loc_cow),  repression = mean(repression), length_camp = n(), success = max(success), regime_support = round(median(regime_support)), 
            camp_support = round(median(camp_support)), camp_size = round(median(camp_size)), pi_armed_wing = mean(pi_armed_wing), pi_police = mean(pi_police), pi_education = mean(pi_education), 
            pi_soc_welfare = mean(pi_soc_welfare), pi_trad_media = mean(pi_trad_media), pi_new_media = mean(pi_new_media), pi_courts = mean(pi_courts),
            colony = max(colony_bool), gdppc = mean(gdppc), prim_meth = mean(prim_meth), sec_defect = mean(sec_defect), xrcomp = mean(xrcomp), in_media = mean(in_media), indiscrim = mean(indiscrim), sdirect = mean(sdirect))

agg_df <- agg_df %>%
  drop_na() %>% 
  mutate(loc_cow = factor(loc_cow)) %>% 
  mutate(end_status = case_when(end_status == 2 ~ 1,
                                TRUE ~ 0))

agg_df %>% 
  ggplot() +
  geom_histogram(aes(x = length_camp), fill = "navy") + 
  labs(
    x = "Length of Campaign in Years",
    title = "Distribution of Campaign Lengths"
  )
  

```

### Longevity Models

```{r}
library(plm)
model1 <- plm(length_camp ~ log(gdppc) + colony + prim_meth + sec_defect + repression + in_media + indiscrim + camp_support + camp_size + regime_support + sdirect + 
                pi_education + pi_police + pi_soc_welfare + pi_trad_media + pi_new_media + pi_courts + sec_defect + 
                prim_meth * pi_education + prim_meth * pi_soc_welfare +
                prim_meth * pi_trad_media + prim_meth * pi_new_media + prim_meth * pi_courts + prim_meth * pi_police, data = agg_df, index = "loc_cow", model="random")

library(RColorBrewer)
nb.cols <- 24
mycolors <- colorRampPalette(brewer.pal(8, "Dark2"))(nb.cols)

broom::tidy(model1, conf.int = TRUE) %>% 
  mutate(term = case_when(term == "sec_defect" ~ "Security Force Defections",
                          term == "repression" ~ "Repression",
                          term == "regime_support" ~ "Regime Support for Campaign", 
                          term == "prim_meth" ~ "Violent Campaign",
                          term == "in_media" ~ "Media Coverage of Campaign",
                          term == "log(gdppc)" ~ "Log(GDP Per Capita)",
                          term == "indiscrim" ~ "Indiscriminate Violence Against Campaign",
                          term == "camp_support" ~ "Campaign Support",
                          term == "camp_size" ~ "Campaign Size",
                          term == "pi_education" ~ "Parallel Education",
                          term == "pi_police" ~ "Parallel Law Enforcement",
                          term == "pi_trad_media" ~ "Parallel Traditional Media",
                          term == "pi_new_media" ~ "Parallel New Media",
                          term == "pi_soc_welfare" ~ "Parallel Social Welfare",
                          term == "pi_courts" ~ "Parallel Judicial System",
                          term == "sdirect" ~ "Sanctions Against Regime",
                          term == "length_camp" ~ "Campaign Length",
                          term == "colony" ~ "Past Colonial History",
                          TRUE ~ term)) %>%
  ggplot(aes(estimate, term, xmin = conf.low, xmax = conf.high, height = 0, color = term)) +
  geom_point() +
  geom_vline(xintercept = 0, lty = 4) +
  geom_errorbarh() +
  guides(color = "none") + 
  labs(x = "Effect on Log Odds of Success",
       y = "Independent Variable",
       title = "Coefficient Plot for Random Effects \n Linear Regression on Length of Campaign") + 
  theme_minimal() + 
  scale_color_manual(values=mycolors)

```


### Binomial Logistic Regression

```{r}
model2 <- glm(end_status ~ length_camp + log(gdppc) + colony + prim_meth + sec_defect + repression + in_media + indiscrim + camp_support + camp_size + regime_support + sdirect + 
                pi_education + pi_police + pi_soc_welfare + pi_trad_media + pi_new_media + pi_courts + sec_defect + prim_meth * pi_education + prim_meth * pi_soc_welfare +
                prim_meth * pi_trad_media + prim_meth * pi_new_media + prim_meth * pi_courts + prim_meth * pi_police, data = agg_df, family = "binomial")

library(RColorBrewer)
nb.cols <- 25
mycolors <- colorRampPalette(brewer.pal(8, "Dark2"))(nb.cols)

broom::tidy(model2, conf.int = TRUE) %>% 
  mutate(term = case_when(term == "sec_defect" ~ "Security Force Defections",
                          term == "repression" ~ "Repression",
                          term == "regime_support" ~ "Regime Support for Campaign", 
                          term == "prim_meth" ~ "Violent Campaign",
                          term == "in_media" ~ "Media Coverage of Campaign",
                          term == "log(gdppc)" ~ "Log(GDP Per Capita)",
                          term == "indiscrim" ~ "Indiscriminate Violence Against Campaign",
                          term == "camp_support" ~ "Campaign Support",
                          term == "camp_size" ~ "Campaign Size",
                          term == "pi_education" ~ "Parallel Education",
                          term == "pi_police" ~ "Parallel Law Enforcement",
                          term == "pi_trad_media" ~ "Parallel Traditional Media",
                          term == "pi_new_media" ~ "Parallel New Media",
                          term == "pi_soc_welfare" ~ "Parallel Social Welfare",
                          term == "pi_courts" ~ "Parallel Judicial System",
                          term == "sdirect" ~ "Sanctions Against Regime",
                          term == "length_camp" ~ "Campaign Length",
                          term == "colony" ~ "Past Colonial History",
                          TRUE ~ term)) %>%
  ggplot(aes(estimate, term, xmin = conf.low, xmax = conf.high, height = 0, color = term)) +
  geom_point() +
  geom_vline(xintercept = 0, lty = 4) +
  geom_errorbarh() +
  guides(color = "none") + 
  labs(x = "Effect on Log Odds of Success",
       y = "Independent Variable",
       title = "Coefficient Plot for Logistic Regression \n on Success of Campaign") + 
  theme_minimal() + 
  scale_color_manual(values=mycolors)
```

### Multinomial Logistic Regression

```{r}
library(nnet)
df$end_status <- as.factor(df$end_status)
df$colony_bool <- as.factor(df$colony_bool)
mn_mod <- multinom(end_status ~ log(gdppc) + colony_bool + years_active + prim_meth + repression + in_media + indiscrim + sdirect +
                     camp_support + camp_size + regime_support + pi_education + pi_police + pi_soc_welfare + pi_trad_media + pi_new_media + pi_courts + sec_defect + 
                     prim_meth * pi_trad_media + prim_meth * pi_new_media + prim_meth * pi_courts + prim_meth * pi_education + prim_meth * pi_police + prim_meth * pi_soc_welfare, data = df, family = "multinomial")



library(interactions)
library(ggstance)
tt <- broom::tidy(mn_mod,conf.int=TRUE)
tt <- tt %>% 
  filter(term!="(Intercept)")  %>% 
  filter(term!="prim_meth1:pi_police1")

tt %>% 
  #filter(!term %in% c("repression", "indiscrim", "sdirect", "camp_support", "years_active", "sec_defect", "regime_support", "in_media", "colony_boolTRUE", "log(gdppc)", "camp_size")) %>%  
  mutate(term = case_when(term == "prim_meth1" ~ "Violent Campaign",
                          term == "pi_education1" ~ "Parallel Education",
                          term == "pi_police1" ~ "Parallel Law Enforcement",
                          term == "pi_trad_media1" ~ "Parallel Traditional Media",
                          term == "pi_new_media1" ~ "Parallel New Media",
                          term == "pi_soc_welfare1" ~ "Parallel Social Welfare",
                          term == "pi_courts1" ~ "Parallel Judicial System",
                          term == "sdirect1" ~ "Sanctions Against Regime",
                          term == "prim_meth1:pi_trad_media1" ~ "Violent Campaign * Traditional Media",
                          term == "prim_meth1:pi_new_media1" ~ "Violent Campaign * New Media",
                          term == "prim_meth1:pi_courts1" ~ "Violent Campaign * Judicial System",
                          term == "prim_meth1:pi_education1" ~ "Violent Campaign * Education",
                          term == "prim_meth1:pi_soc_welfare1" ~ "Violent Campaign * Social Welfare",
                          term == "prim_meth1:pi_police1" ~ "Violent Campaign * Law Enforcement",
                          TRUE ~ term)) %>%
  mutate(y.level = case_when(y.level == 2 ~ "Success",
                             y.level == 1 ~ "Failure")) %>%
  ggplot(aes(estimate, term, xmin = conf.low, xmax = conf.high, height = 0, color = y.level)) +
  ggstance::geom_pointrangeh(position=position_dodgev(height=0.75)) +
  geom_vline(xintercept = 0, lty = 4) +
  ggstance::geom_errorbarh(position=position_dodgev(height=0.75)) +
  labs(x = "Effect on Outcome Rate",
       y = "Independent Variable",
       title = "Coefficient Plot for Multinomial Logit/Competing Risks",
       color = "Outcome") + 
  theme_minimal() 
```
  
### Results

#### Campaign Length

* No independent variable of interest has a clear impact on campaign length; educational institutions _may_ have a slight conditional effect with the violence of the campaign.
* Media coverage shortens campaigns.

#### Campaign Success 

* Across all model specifications, education and social welfare have the strongest effects on success rates. 
* Both competing risks and multinomial logit show that violent campaigns are harmed by building social welfare systems and helped by building educational systems.
  * These results are significant at least at the .1 level in the competing risks regression. 
* I replicate previous findings that nonviolent tactics, security force defections, campaign size, and campaign support from the regime have significant positive effects on success rates. International support of the regime and repression of the campaign have significant negative impacts.

